### 2021-01-29 paper/appendix.tex
Updates to address and read-read

\documentclass[conference,compsoc]{IEEEtran}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Missing includes
\usepackage{graphicx}
\usepackage[prologue]{xcolor}
\definecolor[named]{ACMBlue}{cmyk}{1,0.1,0,0.1}
\definecolor[named]{ACMYellow}{cmyk}{0,0.16,1,0}
\definecolor[named]{ACMOrange}{cmyk}{0,0.42,1,0.01}
\definecolor[named]{ACMRed}{cmyk}{0,0.90,0.86,0}
\definecolor[named]{ACMLightBlue}{cmyk}{0.49,0.01,0,0}
\definecolor[named]{ACMGreen}{cmyk}{0.20,0,1,0.19}
\definecolor[named]{ACMPurple}{cmyk}{0.55,1,0,0.15}
\definecolor[named]{ACMDarkBlue}{cmyk}{1,0.58,0,0.21}
\usepackage[bookmarksnumbered,unicode]{hyperref}
\hypersetup{colorlinks,
  linkcolor=ACMPurple,
  citecolor=ACMPurple,
  urlcolor=ACMDarkBlue,
  filecolor=ACMDarkBlue}
\usepackage{amsmath}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
%\newtheorem{conjecture}{Conjecture}
\newtheorem{proposition}{Prop}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Def}
\newtheorem{example}[definition]{Ex}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \author{\IEEEauthorblockN{%
%     Alan Jeffrey\IEEEauthorrefmark{1} and
%     James Riely\IEEEauthorrefmark{2}
% }
% \IEEEauthorblockA{\IEEEauthorrefmark{1}Roblox}
% \IEEEauthorblockA{\IEEEauthorrefmark{2}DePaul University}}
\author{Anonymous\\Anonymous Institution}

\title{Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers}

\usepackage{macros}
\usepackage[sectionbib,numbers,square]{natbib}

\usepackage{xr}
\newcounter{pomsetPreParCount}\setcounter{pomsetPreParCount}{8}
\newcounter{pomsetXSemiCount}\setcounter{pomsetXSemiCount}{8}
\externaldocument{paper}
\begin{document} 
\pagestyle{plain}
\appendices
\input{discussion}
\bibliographystyle{abbrvnat}
\bibliography{bib}

\end{document}

% Local Variables:
% mode: latex
% TeX-master: t
% End: