### 2021-02-11 paper/appendix.tex
Sorted out bib

\documentclass[conference,compsoc]{IEEEtran}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Missing includes
\usepackage{graphicx}
\usepackage[prologue]{xcolor}
\definecolor[named]{ACMBlue}{cmyk}{1,0.1,0,0.1}
\definecolor[named]{ACMYellow}{cmyk}{0,0.16,1,0}
\definecolor[named]{ACMOrange}{cmyk}{0,0.42,1,0.01}
\definecolor[named]{ACMRed}{cmyk}{0,0.90,0.86,0}
\definecolor[named]{ACMLightBlue}{cmyk}{0.49,0.01,0,0}
\definecolor[named]{ACMGreen}{cmyk}{0.20,0,1,0.19}
\definecolor[named]{ACMPurple}{cmyk}{0.55,1,0,0.15}
\definecolor[named]{ACMDarkBlue}{cmyk}{1,0.58,0,0.21}
\usepackage[bookmarksnumbered,unicode]{hyperref}
\hypersetup{colorlinks,
  linkcolor=ACMPurple,
  citecolor=ACMPurple,
  urlcolor=ACMDarkBlue,
  filecolor=ACMDarkBlue}
\usepackage{amsmath}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Def}
\newtheorem{example}[definition]{Ex}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
%\newtheorem{conjecture}{Conjecture}
\newtheorem{proposition}[definition]{Prop}
\newtheorem{lemma}{Lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \author{\IEEEauthorblockN{%
%     Alan Jeffrey\IEEEauthorrefmark{1} and
%     James Riely\IEEEauthorrefmark{2}
% }
% \IEEEauthorblockA{\IEEEauthorrefmark{1}Roblox}
% \IEEEauthorblockA{\IEEEauthorrefmark{2}DePaul University}}
\author{Anonymous\\Anonymous Institution}

\title{Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers}

\usepackage{macros}
\usepackage[sectionbib,numbers,square]{natbib}

\usepackage{xr,refcount}
\externaldocument{paper}
\setcounter{definition}{\getrefnumber{finaldef}}
\addtocounter{definition}{-1}
\setcounter{figure}{\getrefnumber{finalfig}}
\addtocounter{figure}{-1}
\newcounter{pomsetPreParCount}\setcounter{pomsetPreParCount}{8}
\newcounter{pomsetXSemiCount}\setcounter{pomsetXSemiCount}{8}

\begin{document} 
\pagestyle{plain}
\appendices
% \begin{figure*}
%   \begin{center}
%     \begin{minipage}{.92\textwidth}
%       \input{fig-full}
%       % \input{fig-full-where}
%       % \input{fig-full-noco}
%     \end{minipage}
%   \end{center}
%   \caption{Full Semantics of Loads, Stores and Threads (See \refdef{def:QS}
%     for $\QS{\aLoc}{\amode}$ and $\QL{\aLoc}{\amode}$ and \refdef{def:DS} for
%     $\DL{\aLoc}{\amode}$ and $\DS{\aLoc}{\amode}$)}
%   \label{fig:full}
% \end{figure*}    
\input{discussion}
% \bibliographystyle{abbrvnat}
% \bibliography{bib}

\end{document}

% Local Variables:
% mode: latex
% TeX-master: t
% End: