### 2021-03-29 paper/1SUB/egc.tex
saved old version

\section{The ARMv8 EGC Model}
Copyright info:
\begin{verbatim}
(*
 * The ARMv8 Application Level Memory Model.
 *
 * See section B2.3 of the ARMv8 ARM:
 *   https://developer.arm.com/docs/ddi0487/latest/arm-architecture-reference-manual-armv8-for-armv8-a-architecture-profile
 *
 * Author: Will Deacon <will.deacon@arm.com>
 * Author: Jade Alglave <jade.alglave@arm.com>
 *
 * Copyright (C) 2016-2020, ARM Ltd.
 * All rights reserved.
 *
 * Redistribution and use in source and binary forms, with or without
 * modification, are permitted provided that the following conditions are
 * met:
 *
 *     * Redistributions of source code must retain the above copyright
 *       notice, this list of conditions and the following disclaimer.
 *     * Redistributions in binary form must reproduce the above copyright
 *       notice, this list of conditions and the following disclaimer in
 *       the documentation and/or other materials provided with the
 *       distribution.
 *     * Neither the name of ARM nor the names of its contributors may be
 *       used to endorse or promote products derived from this software
 *       without specific prior written permission.
 *
 * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS
 * IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED
 * TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A
 * PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
 * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
 * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED
 * TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR
 * PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
 * LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING
 * NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS
 * SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 *)

\end{verbatim}

\texttt{armv8-util.cat}:
\begin{verbatim}
"ARMv8 AArch64"

(*
 * Include the cos.cat file shipped with herd.
 * This builds the co relation as a total order over writes to the same
 * location and then consequently defines the fr relation using co and
 * rf.
 *)
include "cos.cat"

(*
 * Include aarch64fences.cat so that barriers show up in generated diagrams.
 *)
include "aarch64fences.cat"

(* Helper functions *)
procedure included(r1, r2) =
  empty r1 \ r2
end

procedure equal(r1, r2) =
  call included(r1, r2)
  call included(r2, r1)
end

let intervening-write(r) = r & (_ * W); r

(*
 * As a restriction of the model, all observers are limited to the same
 * inner-shareable domain. Consequently, the ISH, OSH and SY barrier
 * options are all equivalent to each other.
 *)
let dsb.full = DSB.ISH | DSB.OSH | DSB.SY
let dsb.ld = DSB.ISHLD | DSB.OSHLD | DSB.LD
let dsb.st = DSB.ISHST | DSB.OSHST | DSB.ST

(*
 * A further restriction is that standard litmus tests are unable to
 * distinguish between DMB and DSB instructions, so the model treats
 * them as equivalent to each other.
 *)
let dmb.full = DMB.ISH | DMB.OSH | DMB.SY | dsb.full
let dmb.ld = DMB.ISHLD | DMB.OSHLD | DMB.LD | dsb.ld
let dmb.st = DMB.ISHST | DMB.OSHST | DMB.ST | dsb.st

(* Flag any use of shareability options, due to the restrictions above. *)
flag ~empty (dmb.full | dmb.ld | dmb.st) \
	    (DMB.SY | DMB.LD | DMB.ST | DSB.SY | DSB.LD | DSB.ST)
	    as Assuming-common-inner-shareable-domain

let IM0 = loc & ((IW * (M\IW)) | ((W\FW) * FW))

let intrinsic = (iico_data | iico_ctrl)+
\end{verbatim}

\texttt{armv8-common.cat}:
\begin{verbatim}
"ARMv8 AArch64"

include "armv8-util.cat"

(* Coherence-after *)
let ca = fr | co

(* Local read successor *)
let lrs = [W]; po-loc \ intervening-write(po-loc); [R]

(* Local write successor *)
let lws = po-loc; [W]

(* Observed-by *)
let obs = rfe | fre | coe

(* Read-modify-write *)
let rmw = lxsx | amo

(* Dependency-ordered-before *)
let dob = addr | data
	| ctrl; [W]
	| (ctrl | (addr; po)); [ISB]; po; [R]
	| addr; po; [W]
	| (addr | data); lrs

(* Atomic-ordered-before *)
let aob = rmw
	| [W & range(rmw)]; lrs; [A | Q]

(* Barrier-ordered-before *)
let bob = po; [dmb.full]; po
        | po; ([A];amo;[L]); po
	| [L]; po; [A]
	| [R]; po; [dmb.ld]; po
	| [A | Q]; po
	| [W]; po; [dmb.st]; po; [W]
	| po; [L]

(* Tag-ordered-before *)
let tob = [R & T]; intrinsic; [M \ T]

(* Locally-ordered-before *)
let rec lob = lws
        | dob
	| aob
	| bob
        | tob
	| lob; lob

(* Internal visibility requirement *)
acyclic po-loc | ca | rf as internal

(* Atomic: Basic LDXR/STXR constraint to forbid intervening writes. *)
empty rmw & (fre; coe) as atomic
\end{verbatim}

\texttt{egc.cat}:
\begin{verbatim}
"Armv8 External global completion requirement, not mixed-size"

include "armv8-common.cat"

let gc-req = (W * _) | (R * _) & ((range(rfe) * _) | (rfi^-1; lob))
let preorder-gcb = IM0 | lob & gc-req

with gcb from linearisations(M, preorder-gcb)
~empty gcb

let gcbl = gcb & loc
let rf-gcb = (W * R) & gcbl \ intervening-write(gcbl)
let co-gcb = (W * W) & gcbl

call equal(rf, rf-gcb)
call equal(co, co-gcb)
\end{verbatim}