### 2021-06-28 paper/mca.tex
Reorg

\section{PwT-MCA}

\input{fig-sem.tex}
See \reffig{fig:seq}.