### 2021-07-02 paper/mca1.tex
finished with c11

\section{\PwTmcaTITLE{1}: Pomsets with Predicate Transformers for MCA (Part 1)}
\label{sec:mca1}
We define two models for \mca{} architectures: \PwTmca{1} and \PwTmca{2}.
We drop the subscript when referring to both models.

\begin{definition}
  \label{def:pwt:mca1}
  A \PwTmca{1} is a \PwT{} (\refdef{def:pomset}) equipped
  with a relation $\rrfx$ such that 
  \begin{enumerate}[,label=(\textsc{m}\arabic*),ref=\textsc{m}\arabic*]
    \setcounter{enumi}{\value{Brf}}
  \item \label{pom-rf} \makecounter{rf} ${\rrfx} \subseteq \aEvs\times\aEvs$
    is an injective relation capturing \emph{reads-from}, such that
    % \end{enumerate}
    % A pomset is a \emph{candidate} if there is an injective relation
    % ${\rrfx} : \aEvs\times\aEvs$, capturing \emph{reads-from}, such that
    \begin{enumerate}
      % \begin{enumerate}[,label=(\textsc{i}\arabic*),ref=\textsc{i}\arabic*]
      % \item \label{rf-injective}
      %   if $\bEv\xrfx\aEv$ and $\cEv\xrfx\aEv$ then $\bEv=\cEv$, that is,
      %   ${\rrfx}$ is injective,
    \item \label{pom-rf-match} if $\bEv\xrfx\aEv$ then
      $\labeling(\bEv) \rmatches \labeling(\aEv)$,
    \item \label{pom-rf-block} if $\bEv\xrfx\aEv$ and
      $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\le\bEv$ or
      $\aEv\le\cEv$,
    \item \label{pom-rf-le} if $\bEv\xrfx\aEv$ then $\bEv\le\aEv$.
    \end{enumerate}
  \end{enumerate}

  A \PwTmca{} is \emph{complete} if %it is a complete \PwT{} (\refdef{def:pomset}) and
  \begin{multicols}{2}
    \begin{enumerate}[,label=(\textsc{c}\arabic*),ref=\textsc{c}\arabic*]
    \item[\eqref{top-kappa}]
      \eqref{top-term}\;
      as in \refdef{def:pomset},
      \setcounter{enumi}{\value{Brf}}
    \item \label{top-rf}
      if $\labeling(\aEv)$ is a $\sread$ then there is some $\bEv\xrfx\aEv$.
    \end{enumerate}
  \end{multicols}
\end{definition}

% To include \RMW{}s, also add
% \begin{enumerate}
% \item[(\textsc{p}10)]\;
%   ${\rrmw}=\PBR{{\rrmw_1}\cup{\rrmw_2}}$,
% \end{enumerate}

\input{fig-mca1.tex}
\input{fig-mca2.tex}

The semantic rules are given in \reffig{fig:mca1}.
We write $\semmca{1}{}$ for the semantic function.
Let
\begin{math}
  \semmca{1}{\aCmd_1 \LPAR[\bThrd] \aCmd_2} = \sLPAR{\semmca{1}{\aCmd_1}}{\semmca{1}{\aCmd_2}}.
\end{math}

% From the data model, we require an additional binary relation 
% ${\rdelays}\subseteq(\Act\times\Act)$.  For the actions in this paper, we
% define this as follows.

% \begin{scope}
% \end{scope}