### 2021-07-04 paper/mca1.tex

\section{\PwTmcaTITLE{1}: Pomsets with Predicate Transformers for MCA (Part 1)}
\label{sec:mca1}
We define two models for \mca{} architectures: \PwTmca{1} and \PwTmca{2}.
We drop the subscript when referring to both models.

Note that reads-from implies order by \eqref{pom-rf-le}.
% \begin{lemma}
%   \label{lem:rf:implies:le}
%   For any $\aPS$ in the range of $\sembase{}$, $\bEv\xrfx\aEv$ implies
%   $\bEv\le\aEv$.

%   \vspace{-.5\baselineskip}
%   \begin{proof}
%     Induction on the definition of $\sembase{}$, using \ref{par-rf-le1}, \ref{par-rf-le2}, and \ref{pom-rf-le}.
%   \end{proof}
% \end{lemma}
In top-level pomsets, every read must have a matching write in $\rrfx$
\eqref{top-rf}.  Together with \ref{pom-rf-match} and \ref{pom-rf-block}, the lemma
guarantees that reads are \emph{fulfilled} at top-level, as in
\cite[\textsection 2.7]{DBLP:journals/pacmpl/JagadeesanJR20}.\footnote{The
  basic model would be the same if we move $\rrfx$ from the model itself to
  be existentially quantified in the definition of top-level pomset, along
  with \ref{pom-rf-match} and \ref{pom-rf-block}.  This was the approach of
  \citeauthor{DBLP:journals/pacmpl/JagadeesanJR20} We include $\rrfx$
  explicitly for use in \textsection\ref{sec:arm2}, where we introduce a
  variant semantics $\frf{\semrr{}}$ where \ref{pom-rf-le} is not required.}

From \refdef{def:actions}, recall that $\aAct \rdelaysdef \bAct$ if
$\aAct\eqreorderco\bAct$ or $\aAct\reorderra\bAct$ or
$\aAct\eqreordersc\bAct$.  \ref{seq-le-delays} guarantees that sequential
order is enforced between conflicting accesses of the same location
($\eqreorderco$), into a release and out of an acquire ($\reorderra$), and
between SC accesses ($\eqreordersc$).  Combined with the fulfillment
requirements (\ref{pom-rf-match}, \ref{pom-rf-block} and \ref{pom-rf-le}),
these ensure coherence, publication, subscription and other idioms.  For
example, consider the following:\footnote{We use different colors for arrows representing order:
  \begin{itemize}  
  \item \makebox{$\bEv\xpo\aEv$} arises from control/data/address \emph{dependency} (\ref{seq-kappa}, definition of $\labelingForm'_2(\bEv)$),
  \item \makebox{$\bEv\xsync\aEv$} arises from $\reorderra$ or $\eqreordersc$ \eqref{seq-le-delays},
  \item \makebox{$\bEv\xwki\aEv$} arises from $\eqreorderco$ \eqref{seq-le-delays},
  \item \makebox{$\bEv\xrf\aEv$} arises from \emph{reads-from} \eqref{pom-rf-match},
  %\item \makebox{$\bEv\xsyncsc\aEv$} arises from \emph{strong fencing} \eqref{cand-lesync-sc},
  \item \makebox{$\bEv\xwk\aEv$} arises from \emph{blocking} \eqref{pom-rf-block}.
  \end{itemize}
  In \textsection\ref{sec:arm2}, it is possible for $\rrf$ to contradict
  $\le$.  In this case, we use a dotted arrow for $\rrf$: $\bEv\xrfint\aEv$
  indicates that $\aEv\le\bEv$.}
\begin{gather*}
  \taglabel{pub}
  \begin{gathered}    
    \PW{x}{0}\SEMI 
    \PW{x}{1}\SEMI \PW[\mREL]{y}{1} \PAR \PR[\mACQ]{y}{r}\SEMI \PR{x}{s}
    \\[-.4ex]
    \nonumber
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wx0}{\DW{x}{0}}{}
        \event{wx1}{\DW{x}{1}}{right=of wx0}
        \event{wy1}{\DW[\mREL]{y}{1}}{right=of wx1}
        \event{ry1}{\DR[\mACQ]{y}{1}}{right=2.5em of wy1}
        \event{rx0}{\DR{x}{0}}{right=of ry1}
        \sync{wx1}{wy1}
        \sync{ry1}{rx0}
        \rf{wy1}{ry1}
        \wk[out=170,in=10]{rx0}{wx1}
        \wki{wx0}{wx1}
      \end{tikzinline}}
  \end{gathered}
\end{gather*}
The execution is disallowed due to the cycle.  All of the order shown is
required at top-level: The intra-thread order comes from \ref{seq-le-delays}:
$\DWP{x}{0}\xwki \DWP{x}{1}$ is required by $\eqreorderco$.
$\DWP{x}{1}\xsync \DWP[\mREL]{y}{1}$ and $\DRP[\mACQ]{y}{1}\xsync \DRP{x}{0}$ are
required by $\reorderra$.  The cross-thread order is required by fulfillment:
\ref{top-rf} requires that all top-level reads are in the image of $\xrfx$.
\ref{pom-rf-match} ensures that $\DWP[\mREL]{y}{1}\xrfx \DRP[\mACQ]{y}{1}$, and
\ref{pom-rf-le} subsequently ensures that
$\DWP[\mREL]{y}{1}\le \DRP[\mACQ]{y}{1}$.  The \emph{antidependency}
$\DRP{x}{0}\xwk \DWP{x}{1}$ is required by \ref{pom-rf-block}.  (Alternatively,
we could have $\DWP{x}{1}\xwk \DWP{x}{0}$, again resulting in a cycle.)

The semantics gives the expected results for store buffering and load buffering,
as well as litmus tests involving fences and SC access.  The model of
coherence is weaker than C11, in order to support common subexpression
elimination, and stronger than Java, in order to support local reasoning
about data races.  See
\cite[\textsection3.1]{DBLP:journals/pacmpl/JagadeesanJR20} for a discussion.

\begin{definition}
  \label{def:pwt:mca1}
  A \PwTmca{1} is a \PwT{} (\refdef{def:pomset}) equipped
  with a relation $\rrfx$ such that 
  \begin{enumerate}[,label=(\textsc{m}\arabic*),ref=\textsc{m}\arabic*]
    \setcounter{enumi}{\value{Brf}}
  \item \label{pom-rf} \makecounter{rf} ${\rrfx} \subseteq \aEvs\times\aEvs$
    is an injective relation capturing \emph{reads-from}, such that
    % \end{enumerate}
    % A pomset is a \emph{candidate} if there is an injective relation
    % ${\rrfx} : \aEvs\times\aEvs$, capturing \emph{reads-from}, such that
    \begin{enumerate}
      % \begin{enumerate}[,label=(\textsc{i}\arabic*),ref=\textsc{i}\arabic*]
      % \item \label{rf-injective}
      %   if $\bEv\xrfx\aEv$ and $\cEv\xrfx\aEv$ then $\bEv=\cEv$, that is,
      %   ${\rrfx}$ is injective,
    \item \label{pom-rf-match} if $\bEv\xrfx\aEv$ then
      $\labeling(\bEv) \rmatches \labeling(\aEv)$,
    \item \label{pom-rf-block} if $\bEv\xrfx\aEv$ and
      $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\le\bEv$ or
      $\aEv\le\cEv$,
    \item \label{pom-rf-le} if $\bEv\xrfx\aEv$ then $\bEv\le\aEv$.
    \end{enumerate}
  \end{enumerate}

  \label{def:pwt:mca:complete}
  A \PwTmca{} is \emph{complete} if %it is a complete \PwT{} (\refdef{def:pomset}) and
  \begin{multicols}{2}
    \begin{enumerate}[,label=(\textsc{c}\arabic*),ref=\textsc{c}\arabic*]
    \item[\eqref{top-kappa}]
      \eqref{top-term}\;
      as in \refdef{def:pomset},
      \setcounter{enumi}{\value{Brf}}
    \item \label{top-rf}
      if $\labeling(\aEv)$ is a $\sread$ then there is some $\bEv\xrfx\aEv$.
    \end{enumerate}
  \end{multicols}
\end{definition}

% To include \RMW{}s, also add
% \begin{enumerate}
% \item[(\textsc{p}10)]\;
%   ${\rrmw}=\PBR{{\rrmw_1}\cup{\rrmw_2}}$,
% \end{enumerate}

\input{fig-mca1.tex}
\input{fig-mca2.tex}

The semantic rules are given in \reffig{fig:mca1}.
We write $\semmca{1}{}$ for the semantic function.
Let
\begin{math}
  \semmca{1}{\aCmd_1 \LPAR[\bThrd] \aCmd_2} = \sLPAR{\semmca{1}{\aCmd_1}}{\semmca{1}{\aCmd_2}}.
\end{math}

% From the data model, we require an additional binary relation 
% ${\rdelays}\subseteq(\Act\times\Act)$.  For the actions in this paper, we
% define this as follows.

% \begin{scope}
% \end{scope}