### 2021-10-07 paper/outro.tex

\section{Conclusions} %Future Work}

This paper is the first to present a direct denotational semantics for
sequential composition in a relaxed memory model that can be
efficiently compiled to modern CPUs. We extract from this model a semantic
dependency relation and use it to build \PwTc, a solution to the
Out-of-Thin-Air problem in \cXI, and \PwTmca{}, a model for Java-like
languages.
% There is, as usual, more research to be done.

We have not treated loops in this model, though we expect that the usual
approach of showing continuity for all the semantic operations with respect
to set inclusion would go through. \citet{DBLP:conf/esop/PaviottiCPWOB20} use
step-indexing to account for loops; a similar approach could be applied here.

\PwTmca{} does not validate access elimination: store-forwarding and
dead-write-removal are unsound.  We expect that these can be validated by
allowing events with different actions to merge. % in some instances.

\PwTmca{1} is a simpler model than \PwTmca{2}, but requires fences on
acquiring reads for \armeight{}.  It would be illuminating to find out what
the performance penalty is for these fences.

% In a similar vein,
\PwT{} does not validate read introduction, whereas speculative operational
semantics do.  Recent work shows a tension between read introduction and
compositional reasoning for temporal safety properties (see
\textsection\ref{sec:promising}).  Nonetheless, read introduction is
ubiquitous in some compilers.  It would be interesting to know if there is a
performance penalty for banning read introduction.

% An earlier version of this paper has been mechanized in Agda; it would be
% reassuring to update the mechanization to bring it in line with the current state.

% We have presented the first model of relaxed memory that treats sequential
% composition as a first-class citizen. The model builds directly on \jjr{}.

% For sequential composition, parallel composition and the conditional, we
% believe that the definition is \emph{natural}, even \emph{canonical}.
% For stores and loads, instead, the definition in \reffig{fig:no-addr} is a
% Frankenstein's monster of features.  This complexity is \emph{essential},
% however, not just an accident of our poor choices.  Relaxed memory models must
% please many audiences: compiler writers want one thing, hardware designers
% another, and programmers yet another still.  The result is inevitably full of
% compromise.

% Given that \emph{complexity} cannot be eliminated from relaxed memory models,
% the best one can do is attempt to understand its causes.  We have broken the
% problem into seven manageable pieces, discussed throughout
% \textsection\ref{sec:q}--\ref{sec:complications}.  \refdef{def:pomsets-arm}
% summarizes all the features necessary for efficient implementation on
% \armeight{}.  We discuss address calculation, read-modify-write operations
% and fences in the appendix.

% {Logic} is the thread that sews these features together.

% %A unique feature of our model is the centrality of logic:

% % study each feature in isolation, as a small delta to the base definition
% % given in \textsection\ref{sec:model}.
% % In we studied eight such
% % features.

% % seem staggeringly complex, we argue that this complexity is unavoidable if
% % one wants all the features that it embodies.  By breaking the definition into
% % its constituent parts, we have shown how each of eight

% \subsection*{Acknowledgements}
% % This paper has been greatly improved by the comments of the anonymous reviewers.
% Riely was supported by the National Science Foundation under
% grant No.~CCR-1617175.

% % It is based upon work supported by the National Science Foundation under
% % Grant No. CCR-1617175. Any opinions, findings, and conclusions or
% % recommendations expressed in this material are those of the author and do
% % not necessarily reflect the views of the NSF.}