### 2021-10-28 paper/outro.tex

\section{Limitations and Future Work}
\label{sec:outro}

This paper is the first to present a direct denotational semantics for
sequential composition in a relaxed memory model that can be efficiently
compiled to modern CPUs. We extract from this model a semantic dependency
relation and use it to build \PwTc, a solution to the Out-of-Thin-Air problem
in \cXI, and \PwTmca{}, a model intended for safe languages such as Java and
Javascript.  Our work has several limitations, providing opportunities for
future work.

We have mechanized some proofs, but not all.  In particular, we have only a
pen-and-paper proof showing that \PwTmca{} supports optimal lowering to
\armeight{}.  The same is true for local data race-freedom (\ldrfsc{}).
Additionally, our proof sketch for \ldrfsc{} elides \RMW{}s, which have
caused complications in other models \cite{promising-ldrf}.

We have not treated loops, although we expect that the usual approach of
showing continuity for all the semantic operations with respect to set
inclusion would go through. \citet{DBLP:conf/esop/PaviottiCPWOB20} use
step-indexing to account for loops; perhaps this approach could be adapted.

\PwTmca{} does not validate access elimination: store-forwarding and
dead-write-removal are unsound.  We expect that these can be validated by
allowing events with different actions to merge. % in some instances.
Nor does \PwTmca{} validate the reverse inclusions for
\reflem{lem:if}\eqref{lem:ifelse:if:if}. The culprit is $\rdelay$, which
introduces order regardless of whether preconditions are disjoint.  As an
example, using augmentation, 
\begin{math}
  \sem{\IF{r}
  \THEN \PW{x}{1}
  \ELSE \PW{x}{2}
  \FI}
\end{math}
has an execution with
\begin{math}
  (r{=}0\mid\DW{x}{2})
  \xwki
  (r{\neq}0\mid\DW{x}{1}),
\end{math}
whereas
\begin{math}
  \sem{
    \IF{r} \THEN \PW{x}{1}\FI
    \SEMI
    \IF{\BANG r} \THEN \PW{x}{2}\FI
  \FI}
\end{math}
has no such execution.  We expect that this can be remedied by encoding
$\rdelay$ in the logic.
%\todo{Comment on \PwTpo{} and \PwTc{}?}

\PwTmca{1} is a simpler model than \PwTmca{2}, but requires fences on
acquiring reads for \armeight{}.  It would be illuminating to find out what
the performance penalty is for these fences.  In a similar vein, it would be
interesting to know if there is a performance penalty for banning
read-introduction, which is not valid for \PwT{}.

\PwTc{} can be lowered efficiently to any architecture supported by \cXI{},
but inherits the top-level axioms of \rcXI{}, compromising compositionality.
\PwTmca{} is as a compositional as a model of concurrent imperative
programming can be, but is limited to \mca{} architectures for optimal
lowering.  It would be interesting to explore the middle ground to find a
fully compositional model that supports optimal lowering to all modern
architectures.

% In a similar vein, \PwT{} does not validate read-introduction, whereas
% speculative operational semantics do.  Recent work shows a tension between
% read-introduction and compositional reasoning for temporal safety properties
% (see \textsection\ref{sec:related}).  Nonetheless, read introduction is
% ubiquitous in some compilers.  It would be interesting to know if there is a
% performance penalty for banning read-introduction.

% An earlier version of this paper has been mechanized in Agda; it would be
% reassuring to update the mechanization to bring it in line with the current state.

% We have presented the first model of relaxed memory that treats sequential
% composition as a first-class citizen. The model builds directly on \jjr{}.

% For sequential composition, parallel composition and the conditional, we
% believe that the definition is \emph{natural}, even \emph{canonical}.
% For stores and loads, instead, the definition in \reffig{fig:no-addr} is a
% Frankenstein's monster of features.  This complexity is \emph{essential},
% however, not just an accident of our poor choices.  Relaxed memory models must
% please many audiences: compiler writers want one thing, hardware designers
% another, and programmers yet another still.  The result is inevitably full of
% compromise.

% Given that \emph{complexity} cannot be eliminated from relaxed memory models,
% the best one can do is attempt to understand its causes.  We have broken the
% problem into seven manageable pieces, discussed throughout
% \textsection\ref{sec:q}--\ref{sec:complications}.  \refdef{def:pomsets-arm}
% summarizes all the features necessary for efficient implementation on
% \armeight{}.  We discuss address calculation, read-modify-write operations
% and fences in the appendix.

% {Logic} is the thread that sews these features together.

% %A unique feature of our model is the centrality of logic:

% % study each feature in isolation, as a small delta to the base definition
% % given in \textsection\ref{sec:model}.
% % In we studied eight such
% % features.

% % seem staggeringly complex, we argue that this complexity is unavoidable if
% % one wants all the features that it embodies.  By breaking the definition into
% % its constituent parts, we have shown how each of eight

Supplementary material for this paper available at
\url{https://weakmemory.github.io/pwt}.

\subsection*{Acknowledgements}
This paper has been greatly improved by the comments of the anonymous reviewers.

James Riely was supported by the National Science Foundation under
grant No.~CCR-1617175.

\todo{Add support for Mark\&Simon.}

Anton Podkopaev was supported by JetBrains Research.

% % It is based upon work supported by the National Science Foundation under
% % Grant No. CCR-1617175. Any opinions, findings, and conclusions or
% % recommendations expressed in this material are those of the author and do
% % not necessarily reflect the views of the NSF.}