### 2021-11-08 paper/appendix.tex
make appendices a combo

\RequirePackage{amssymb}  %% bizarre error previous def of Bbbk if after documentclass
\documentclass[acmsmall,screen]{acmart}\settopmatter{printfolios=true}
\hypersetup{bookmarksnumbered,bookmarksopen=true,bookmarksdepth=3}
\settopmatter{printfolios=true}
\AtEndPreamble{%
  \theoremstyle{acmdefinition}
  \newtheorem{remark}[theorem]{Remark}
  \newtheorem{candidate}[theorem]{Candidate}
  \renewcommand{\theequation}{\fnsymbol{equation}}
}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations

\setcopyright{rightsretained}
\acmPrice{}
\acmDOI{}
\acmYear{2022}
\copyrightyear{2022}
\acmSubmissionID{}
\acmJournal{PACMPL}
\acmVolume{0}
\acmNumber{POPL}
\acmArticle{0}
\acmMonth{1}
\startPage{1}

\usepackage{macros}
% \showRAfalse
\showSCOPEfalse
\title{The Leaky Semicolon (Supplementary Material)}

\usepackage{xr}
\externaldocument{paper}
\makecounter{Bkappa} \setcounter{Bkappa}{2}
\makecounter{Btau}   \setcounter{Btau}{3}
\makecounter{Bterm}  \setcounter{Bterm}{4}
\begin{document}
\setcounter{page}{31}
\appendix
\noindent
{\LARGE \textsf{\textbf{The Leaky Semicolon (Supplementary Material)}}}
\bigskip

\input{arm.tex}
\input{oopsla.tex}
\input{discussion.tex}
\input{extra.tex}
\input{drf.tex}
\bibliography{bib}
\end{document}

% Local Variables:
% mode: latex
% TeX-master: t
% End: